---
title: "DBC Style Guide Usage and Examples"
author: "Daniel Carpenter"
format: 
    gfm:
      toc: true
      toc-depth: 3
      number-sections: true
editor: visual
---

# Installation

```{r, eval=FALSE}
# install.packages("devtools")
remotes::install_github("Daniel-Carpenter/styles")
```


```{r, echo=FALSE, message=FALSE}
devtools::load_all(".")
```


# Basic Usage of Functions

## Get the Theme

```{r}
# Retreive theme with helpful theme functions
library(styles)
```

## Colors

### What Colors are Available?

-   Notice `displayNames = TRUE`, which show you what hex codes are associated with the color names.

-   Default is `displayNames = FALSE` for best functionality with plotting

#### Fill colors

```{r}
scale_dbc('fill', displayNames = TRUE)
```

#### Line colors

```{r}
scale_dbc('color', displayNames = TRUE)
```

#### Base colors, like the ones on the navy background of slide decks

```{r}
scale_dbc('base', displayNames = TRUE)
```

#### Text, Background, Border, and Greyed out colors

-   These are all separate since usually only selecting one

```{r}
scale_dbc('text')       # Text (dark gray) 
scale_dbc('background') # Background (white)
scale_dbc('border')     # Border (light gray)
scale_dbc('grayedOut')  # (Very light gray)
```

<br>

### How to get `1` color?

```{r}
# Single color from the fill palette
scale_dbc('fill', 'red')

# Single color from the base palette
scale_dbc('base', 'navy')
```

<br>

### How to get `>1` colors

```{r}
# Get the first 3 colors in the line palette
scale_dbc('color', 3)

# Get the last 3 colors in the fill palette
scale_dbc('color')[6:8]

# Or access specific colors all at once
scale_dbc('color', 'blue', 'orange', 'green', 'yellow')
```

<br>

------------------------------------------------------------------------

<br>

# `ggplot2` Integration

### Create a basic plot structure

> Does *NOT* include DBC Styles

```{r}
basePlot <- ggplot(data = mtcars,
               aes(x     = wt,
                   shape = as.factor(cyl),
                   color = as.factor(cyl),
                   y     = mpg
                   )
               ) +
          geom_point(size = 5)
basePlot # Display
```

<br>

## Use Multiple Colors

> Using DBC Style

```{r}
basePlot +
  
  # Use the DBC Fill palette
  scale_color_dbc() +
  
  # Get the DBC ggplot theme
  theme_dbc()
```

<br>

## Use only `1` Color

> Using DBC Style

```{r}
basePlot + 
  
  # Use only a single color (note using line palette)
  geom_point(color = scale_dbc('color', 'blue'),
             size  = 5) +
  
  # Get the DBC ggplot theme
  theme_dbc()
```
