---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# `styles`

<!-- badges: start -->

<!-- badges: end -->

## Overview

> -   Brand in `ggplot2`
> -   Package that streamlines color palette usage, initially adapted from the [Tableau 10 Palette](https://www.tableau.com/blog/colors-upgrade-tableau-10-56782)
> -   See colorblind statistics for the [fill palette](https://davidmathlogic.com/colorblind/#%23BECDE0-%23FFD39F-%23FFB5B3-%23B4DFCE-%23CDC9E0-%23F2D4A0-%23C3C6D2-%23E6D7D7) and the [color (*line*) palette](https://davidmathlogic.com/colorblind/#%237996B9-%23DBA657-%23C86D6A-%2383AF9E-%23928EA4-%23C2A570-%23828591-%23BAABAB)
> -   ***PLEASE SEE THE PACKAGE [VIGNETTE]{.underline} FOR MORE DETAILED EXAMPLES***

<br>

## Installation of Package

You can install the development version of styles from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# Install devtools if not installed (for GitHub Package)
if (!require("devtools"))   install.packages("devtools")

# Install the styles repository
remotes::install_github("Daniel-Carpenter/styles", build_vignettes = TRUE)
```

## Viewing the Vignette

To see in-depth documentation of the package, please uncomment and issue the following commands.

```{r, eval=FALSE}
# See what vignettes are available
browseVignettes('styles')
```

## Examples

```{r}
library(styles)
```

### Create a basic plot structure

> Plot that does *NOT* include Styles

```{r}
library(ggplot2)

basePlot <- ggplot(data = mtcars,
                   aes(x     = wt,
                       shape = as.factor(cyl),
                       color = as.factor(cyl),
                       y     = mpg
                   )
) +
  geom_point(size = 5)
basePlot # Display
```

### Use Multiple Colors

#### Default Color Palette

> Using Style

```{r}
basePlot +
  
  # Use the color palette
  scale_color_dc() +
  
  # Get the ggplot theme
  theme_dc()
```

#### Override Color with Fill

> Using Style

```{r}
basePlot +
  
  # Use the Fill palette (override)
  scale_color_dc(overrideWithFill = TRUE) +  
  
  # Get the ggplot theme
  theme_dc()
```

### Use only `1` Color

> Using Style

```{r}
basePlot + 
  
  # Use only a single color (note using line palette)
  geom_point(color = scale_dc('base', 'base2'),
             size  = 5) +
  
  # Get the ggplot theme
  theme_dc()
```

### How to use Times New Roman

> Use this setting for ***non-website*** reporting.

#### Non-Label Font Control

```{r}
basePlot +
  
  # Use the color palette
  scale_color_dc() +
  
  # Get the ggplot theme
  theme_dc(inTimesNewRoman = TRUE)
```

#### Label Font Control

```{r}
basePlot + 
  
  geom_text(aes(label = paste(mpg)),
            vjust = 1.2, hjust = -0.2,
            size = 3.5,
            
            # ** IMPORTANT: theme() does not control text labels
            # Need to set family to this font!
            family = styles::getTimesNewRomanFamily()
  ) + 
  
  # This controls all non-label fonts
  theme_dc(inTimesNewRoman = TRUE)
```

### Darken Text when Over Filled Elements

```{r}
ggplot(mtcars, aes(y = mpg, x = wt, color = as.factor(cyl))) +
  geom_point(aes(fill = as.factor(cyl)),
             size = 10,
             pch = 21,
             color = 'transparent',
             alpha = 0.9
  ) +
  geom_text(aes(label = round(mpg, 0))) +
  scale_fill_dc() +
  
  # KEY - darken the text so that it is easier to view
  # Over fill
  scale_color_dc(darkenPaletteForTextGeoms = TRUE) +
  theme_dc()
```

### Offset or Reverse Order of Colors

```{r}
# Offset the colors by 1
basePlot + scale_color_dc(colorOffset = 5)

# reverse the order of the palette
basePlot + scale_color_dc(reverseOrder = TRUE)
```

### Override the palette with a color blind palette

```{r}
# Use color blind friendly palette (works with fill too)
basePlot + scale_color_dc(useColorBlindPalette = TRUE)

# Change the palette (can use cols4all::c4a_palettes() to try others)
# Also, can demo others in GUI using cols4all::c4a()
basePlot + scale_color_dc(useColorBlindPalette = TRUE, 
                          colorBlindPaletteName = 'color_blind')
```

## Numeric Formats

> Mainly for quick financial axis formatting

```{r}
kDollarsFormat(1000,          scaleUnit = 'K')
kDollarsFormat(1000000,       scaleUnit = 'M')
kDollarsFormat(1000000,       scaleUnit = 'MM')
kDollarsFormat(1000000000,    scaleUnit = 'B')
kDollarsFormat(1500000000000, scaleUnit = 'T')
kDollarsFormat(1000000,       scaleUnit = 'M', useDollarSign = FALSE)
```

## Colors Deep Dive

### What Colors are Available?

-   Notice `displayNames = TRUE`, which show you what hex codes are associated with the color names.

-   Default is `displayNames = FALSE` for best functionality with plotting

#### Fill colors

```{r}
scale_dc('fill', displayNames = TRUE)
```

#### Accent colors

```{r}
scale_dc('color', displayNames = TRUE)
```

#### Blue and Gray colors, like the background of slide decks

```{r}
scale_dc('base', displayNames = TRUE)
scale_dc('gray', displayNames = TRUE)
```

#### Text, Grays, and Blues

-   Common colors that you might need

```{r}
scale_dc('text') # Text (dark gray) 
scale_dc('base') # Grays that are in the brand
scale_dc('gray') # Blues that are in the brand
```

### How to get `1` color?

```{r}
# Single color from the fill palette
scale_dc('fill', 'red')

# Single color from the base palette
scale_dc('gray', 'gray3')

# Single color from the base palette
scale_dc('base', 'base2')
```

### How to get `>1` colors

```{r}
# Get the first 3 colors in the line palette
scale_dc('color', 3)

# Get the last 3 colors in the fill palette
scale_dc('color')[6:8]

# Or access specific colors all at once
scale_dc('color', 'base', 'orange', 'green', 'yellow')
```
