---
title: "DBC Style Guide Usage and Examples"
author: "Daniel Carpenter"
format: 
    gfm:
      toc: true
      toc-depth: 3
      number-sections: true
editor: visual
---

<br>

# Overview
> * Package that streamlines color palette usage, adapated from the [Tableau 10 Palette](https://www.tableau.com/blog/colors-upgrade-tableau-10-56782)  
> * Leverages `ggplot2`-like functionality  
> * See colorblind statistics for the [fill palette](https://davidmathlogic.com/colorblind/#%23BECDE0-%23FFD597-%23F6B7B4-%23BEE0D2-%23E4C6DC-%23BDDBE1-%23F7E5B3-%23DCDADA-%23FCC8DA-%23E5CFC5) and the [color (*line*) palette](https://davidmathlogic.com/colorblind/#%236388B4-%23E68900-%23EB4B43-%2354AC88-%23B07AA1-%233C9DAA-%23C5A952-%238A807E-%23D16D91-%23945430)


<br>

# Installation of Package

```{r, eval=FALSE}
# Install devtools if not installed (for GitHub Package)
if (!require("devtools"))   install.packages("devtools")

# Install the styles repository
remotes::install_github("Daniel-Carpenter/styles")
```



# `ggplot2` Integration

```{r}
# Retrieve theme with helpful theme functions
library(styles)
library(ggplot2)
```

### Create a basic plot structure
> Does *NOT* include DBC Styles

```{r}
basePlot <- ggplot(data = mtcars,
               aes(x     = wt,
                   shape = as.factor(cyl),
                   color = as.factor(cyl),
                   y     = mpg
                   )
               ) +
          geom_point(size = 5)
basePlot # Display
```

<br>

## Use Multiple Colors

### Default Color Palette

> Using DBC Style

```{r}
basePlot +
  
  # Use the DBC color palette
  scale_color_dbc() +
  
  # Get the DBC ggplot theme
  theme_dbc()
```

<br>

### Override Color with Fill

> Using DBC Style

```{r}
basePlot +
  
  # Use the DBC Fill palette (override)
  scale_color_manual(values = scale_dbc('fill', 3) ) +
  
  # Get the DBC ggplot theme
  theme_dbc()
```

<br>

## Use only `1` Color

> Using DBC Style

```{r}
basePlot + 
  
  # Use only a single color (note using line palette)
  geom_point(color = scale_dbc('color', 'blue'),
             size  = 5) +
  
  # Get the DBC ggplot theme
  theme_dbc()
```

<br>

# Colors Deep Dive

## What Colors are Available?

-   Notice `displayNames = TRUE`, which show you what hex codes are associated with the color names.

-   Default is `displayNames = FALSE` for best functionality with plotting

### Fill colors

```{r}
scale_dbc('fill', displayNames = TRUE)
```

### Line colors

```{r}
scale_dbc('color', displayNames = TRUE)
```

### Base colors, like the ones on the navy background of slide decks

```{r}
scale_dbc('base', displayNames = TRUE)
```

### Text, Background, Border, and Greyed out colors

-   These are all separate since usually only selecting one

```{r}
scale_dbc('text')       # Text (dark gray) 
scale_dbc('background') # Background (white)
scale_dbc('border')     # Border (light gray)
scale_dbc('grayedOut')  # (Very light gray)
```

<br>

## How to get `1` color?

```{r}
# Single color from the fill palette
scale_dbc('fill', 'red')

# Single color from the base palette
scale_dbc('base', 'navy')
```

<br>

## How to get `>1` colors

```{r}
# Get the first 3 colors in the line palette
scale_dbc('color', 3)

# Get the last 3 colors in the fill palette
scale_dbc('color')[6:8]

# Or access specific colors all at once
scale_dbc('color', 'blue', 'orange', 'green', 'yellow')
```

<br>

------------------------------------------------------------------------

